Optimal (Multiway) Spatial Joins

### 2024-06-03 msj.tex
yufei

\extraspacing \underline{\em Proof of Statement (3).} We will prove 
\myeqn{
    \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')} |\dcross_H(\bm{t})| \cdot |\dcross_V(\bm{t})| \leq \out. \label{eqn:hv:type1:properties:proof-help1}
} 
which implies statement (3) because, for any $\bm{t}$ in the summation, $|\dcross_H(\bm{t})| \ge 1$ and $|\dcross_V(\bm{t})| \ge 1$ by Proposition~\ref{prop:app:hv:type1:help2}. Our proof resorts to the algorithm generate-$\J^*$ (see Section~\ref{sec:hv:type1}). This algorithm adds to $\J^*$ exactly as many tuples as calculated by the left hand side of \eqref{eqn:hv:type1:properties:proof-help1}. By statement (1) of Lemma~\ref{lmm:hv:type2:properties}, the $\J^*$ produced must be a subset of $\J(R_1,...,R_{k-2},H,V)$, whereas $\out = |\J(R_1,...,R_{k-2},H,V)|$. This establishes the inequality in \eqref{eqn:hv:type1:properties:proof-help1}, which completes the proof of statement (3).

@@

\extraspacing \underline{\em Proof of Statement (1).} We will map each $r^*$ of $\bigcup_{i = 1}^{k-2} R_i^*$ to a unique tuple $\bm{t}$ in $\J(R_1',...,R_{k-2}')$ satisfying $\dcross_H(\bm{t})$ $= \cross_H(r^*)$. After this, we can then derive
\myeqn{
    \sum_{i\in[k-2]}\sum_{r^*\in R_i^*} |\cross_H(r^*)| \leq \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}|\dcross_H(\bm{t})|\leq \out \nn
}
where the last step used  statement (3) of Lemma~\ref{lmm:hv:type1:properties}.

@@

The mapping is designed as follows. Consider an arbitrary $r^*\in \bigcup_{i = 1}^{k-2} R_i^*$. Recall that $r^*$ is the top-sliced of some rectangle $r'$. Specifically, if $r' = [x_1,x_2]\times [y_1,y_2]$, then $r^* = [x_1,x_2]\times [y_1,\maxtop(r')]$. By the definition of $\maxtop(r')$ in \eqref{eqn:hv:type1:topsliced}, there exists a tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ satisfying $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_{\bm{t}})$. We map $r^*$ to $\bm{t}$.

@@

Next, we will prove that a segment $h \in \cross_H(r^*)$ if and only if $h \in \dcross_H(\bm{t})$. Let us start with the ``only-if direction'': assuming $h \in \cross_H(r^*)$, we aim to show $h \in \dcross_H(\bm{t})$, or in other words, $h$ crosses both $\gbot(\bm{t}) = r'$ and $B_\bm{t}$.
\myitems {
    \item First, we explain why $h$ crosses $r'$. Because $r^* \subseteq r'$, the fact of $h$ crossing $r^*$ tells us that $h \cap r' \ne \emptyset$. On the other hand, the fact of $h$ crossing $r^*$ also tells us that $\xleft(h) < x_1 \le x_2 < \xright(h)$ (recall that $[x_1, x_2]$ is the x-range of $r^*$). As $[x_1, x_2]$ is also the x-range of $r'$, the fact of $h \cap r' \ne \emptyset$ indicates that $h$ must cross $r'$.

@@

    \item Then, we explain why $h$ crosses $B_{\bm{t}}$. Because $B_{\bm{t}}\subseteq r'$ and $h$ crosses $r'$ (as was just proved), the rectangle $B_{\bm{t}}$ cannot contain any endpoint of $h$. Let us take the y-coordinate $y_h$ of $h$. We will prove that the point $p = (\xleft(B_{\bm{t}}), y_h)$ is on the segment $h \cap B_{\bm{t}}$, suggesting $h\cap B_{\bm{t}} \neq \emptyset$. Once done, this will indicate that $h$ crosses $B_{\bm{t}}$ (because no endpoint of $h$ falls in $B_\bm{t}$).

@@

    As $\gbot(\bm{t}) = r'$, the bottom edge of $B_{\bm{t}}$ is contained in $r'$. Thus, $\xleft(B_{\bm{t}})\in [x_1,x_2]$. As proved earlier, $\xleft(h) < x_1 \le x_2 < \xright(h)$, which leads to $\xleft(B_{\bm{t}}) \in [\xleft(h), \xright(h)]$, suggesting that $p \in h$.

@@

    As $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_\bm{t})$, the y-range of $B_\bm{t}$ is $[y_1, \maxtop(r')]$. Because $h \cap r^* \ne \emptyset$ (as $h$ crosses $r^*$) and $[y_1, \maxtop(r')]$ is the y-range of $r^*$, we know that $y_h \in [y_1, \maxtop(r')]$, suggesting that $p \in B_{\bm{t}}$.
}

We now proceed to establish the ``if direction'': assuming $h \in \dcross_H(\bm{t})$, we aim to show $h \in \cross_H(r^*)$. Recall that $r^* = [x_1,x_2] \times [y_1, \maxtop(r')]$. To prove that $h$ crosses $r^*$, we will argue that $\xleft(h) < x_1 \le x_2 < \xright(h)$ and $y_h \in [y_1, \maxtop(r')]$, where $y_h$ is the y-coordinate of $h$.
\myitems{
    \item The fact $h \in \dcross_H(\bm{t})$ tells us that $h$ crosses $r'$. As $[x_1, x_2]$ is the x-range of $r'$, we know $\xleft(h) < x_1 \le x_2 < \xright(h)$.

    \item The fact $h \in \dcross_H(\bm{t})$ also tells us that $h$ crosses $B_{\bm{t}}$. As $[y_1, \maxtop(r')]$ is the y-range of $B_\bm{t}$ (because $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_\bm{t})$), we know $y_h \in [y_1, \maxtop(r')]$.
}

>>>>

\extraspacing \underline{\em Proof of Statement (3).} We will prove 
\myeqn{
    \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')} |\dcross_H(\bm{t})| \cdot |\dcross_V(\bm{t})| \leq \out \label{eqn:hv:type1:properties:proof-help1}
} 
which implies statement (3) because, for any $\bm{t}$ in the summation, $|\dcross_H(\bm{t})| \ge 1$ and $|\dcross_V(\bm{t})| \ge 1$ by Proposition~\ref{prop:app:hv:type1:help2}. Our proof resorts to the algorithm generate-$\J^*$ (see Section~\ref{sec:hv:type1}). This algorithm adds to $\J^*$ exactly as many tuples as calculated by the left hand side of \eqref{eqn:hv:type1:properties:proof-help1}. By statement (1) of Lemma~\ref{lmm:hv:type2:properties}, the $\J^*$ produced must be a subset of $\J(R_1,...,R_{k-2},H,V)$, whereas $\out = |\J(R_1,...,R_{k-2},H,V)|$. This establishes the inequality in \eqref{eqn:hv:type1:properties:proof-help1}, which completes the proof of statement (3).

@@

\extraspacing \underline{\em Proof of Statement (1).} We will map each $r^*$ of $\bigcup_{i = 1}^{k-2} R_i^*$ to a unique tuple $\bm{t}$ in $\J(R_1',...,R_{k-2}')$ satisfying $\cross_H(r^*) \subseteq \dcross_H(\bm{t})$. The mapping allows us to derive
\myeqn{
    \sum_{i\in[k-2]}\sum_{r^*\in R_i^*} |\cross_H(r^*)| \leq \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}|\dcross_H(\bm{t})|\leq \out \nn
}
where the last step used  statement (3) of Lemma~\ref{lmm:hv:type1:properties}.

@@

The mapping is as follows. Consider an arbitrary $r^*\in \bigcup_{i = 1}^{k-2} R_i^*$. Recall that $r^*$ is the top-sliced of some rectangle $r'$. Specifically, if $r' = [x_1,x_2]\times [y_1,y_2]$, then $r^* = [x_1,x_2]\times [y_1,\maxtop(r')]$. By the definition of $\maxtop(r')$ in \eqref{eqn:hv:type1:topsliced}, there exists a tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ satisfying $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_{\bm{t}})$. We map $r^*$ to $\bm{t}$.

@@

Next, we will prove 
\begin{center}
    {\bf Claim 1:} If a segment $h \in \cross_H(r^*)$, then $h \in \dcross_H(\bm{t})$.
\end{center}
To prove the claim, we must show $h \in \dcross_H(\bm{t})$, or in other words, $h$ crosses both $\gbot(\bm{t}) = r'$ and $B_\bm{t}$.
\myitems {
    \item First, we explain why $h$ crosses $r'$. Because $r^* \subseteq r'$, the fact of $h$ crossing $r^*$ tells us that $h \cap r' \ne \emptyset$. On the other hand, the fact of $h$ crossing $r^*$ also tells us that $\xleft(h) < x_1 \le x_2 < \xright(h)$ (recall that $[x_1, x_2]$ is the x-range of $r^*$). As $[x_1, x_2]$ is also the x-range of $r'$, the fact of $h \cap r' \ne \emptyset$ indicates that $h$ must cross $r'$.

@@

    \item Then, we explain why $h$ crosses $B_{\bm{t}}$. Because $B_{\bm{t}}\subseteq r'$ and $h$ crosses $r'$ (just proved), the rectangle $B_{\bm{t}}$ cannot contain any endpoint of $h$. Let us take the y-coordinate $y_h$ of $h$. We will prove that the point $p = (\xleft(B_{\bm{t}}), y_h)$ is on the segment $h \cap B_{\bm{t}}$, suggesting $h\cap B_{\bm{t}} \neq \emptyset$. Once done, this will indicate that $h$ crosses $B_{\bm{t}}$ (because no endpoint of $h$ falls in $B_\bm{t}$).

@@

    \myitems{
        \item As $\gbot(\bm{t}) = r'$, the bottom edge of $B_{\bm{t}}$ is contained in $r'$. Thus, $\xleft(B_{\bm{t}})\in [x_1,x_2]$. As proved earlier, $\xleft(h) < x_1 \le x_2 < \xright(h)$, which leads to $\xleft(B_{\bm{t}}) \in [\xleft(h), \xright(h)]$, suggesting that $p \in h$.

@@

        \item As $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_\bm{t})$, the y-range of $B_\bm{t}$ is $[y_1, \maxtop(r')]$. Because $h \cap r^* \ne \emptyset$ (as $h$ crosses $r^*$) and $[y_1, \maxtop(r')]$ is the y-range of $r^*$, we know that $y_h \in [y_1, \maxtop(r')]$, suggesting that $p \in B_{\bm{t}}$.
    }
}

% We now turn to the ``if direction'': assuming $h \in \dcross_H(\bm{t})$, we aim to show $h \in \cross_H(r^*)$. Recall that $r^* = [x_1,x_2] \times [y_1, \maxtop(r')]$. To prove that $h$ crosses $r^*$, we will argue that $\xleft(h) < x_1 \le x_2 < \xright(h)$ and $y_h \in [y_1, \maxtop(r')]$, where $y_h$ is the y-coordinate of $h$.
% \myitems{
%     \item The fact $h \in \dcross_H(\bm{t})$ tells us that $h$ crosses $r'$. As $[x_1, x_2]$ is the x-range of $r'$, we have $\xleft(h) < x_1 \le x_2 < \xright(h)$.
% 
%     \vgap
% 
%     \item The fact $h \in \dcross_H(\bm{t})$ also tells us that $h$ crosses $B_{\bm{t}}$. As $[y_1, \maxtop(r')]$ is the y-range of $B_\bm{t}$ (because $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_\bm{t})$), we have $y_h \in [y_1, \maxtop(r')]$.
% }
